---
title: "Configuration Guide for appendMCP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuration Guide for appendMCP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(appendMCP)
library(dplyr)
library(knitr)
library(purrr)
```

## Introduction

This vignette provides a comprehensive guide to configuring studies in the `appendMCP` package. We'll examine each component of the configuration object using the built-in example study to understand the structure and meaning of all data elements.

## Configuration Overview

A complete study configuration consists of 9 main components:

```{r load-config}
# Load the example configuration
data("example_study_config")
config <- example_study_config

# Display the top-level structure
names(config)
```

Let's examine each component in detail:

## 1. Study Metadata

### Study Name and Description

```{r metadata}
cat("Study Name:", config$study_name, "\n")
cat("Description:", config$study_description, "\n")
cat("Alpha Level:", config$alpha, "\n")
```

**Fields:**
- `study_name`: Character string identifying the study
- `study_description`: Brief description of the study design
- `alpha`: Overall Type I error rate (typically 0.025 for one-sided tests)

## 2. Analyses Specification

The `analyses` data frame defines when analyses will be conducted. Due to list columns, we display it in parts:

```{r analyses-part1}
# Part 1: Basic trigger columns
kable(config$analyses[,c("endpoint", "strata", "treatments", "sample_size", "events")],
      caption = "Analyses Specification - Part 1: Trigger Conditions")
```

```{r analyses-part2}
# Part 2: Power subset specifications (displayed as text due to nested lists)
cat("\nPart 2: Power Subset Specifications\n")
cat("===================================\n")
for(i in seq_len(nrow(config$analyses))) {
  cat("\nAnalysis", i, "(endpoint:", config$analyses$endpoint[i], "):\n")
  cat("  power_subsets_any:", paste(names(config$analyses$power_subsets_any[[i]]), collapse = ", "), "\n")
  cat("  power_subsets_all:", paste(names(config$analyses$power_subsets_all[[i]]), collapse = ", "), "\n")
}
```

**Row Interpretation:**

- Each row represents one planned analysis and describes the conditions for an analysis to occur
- Analysis timing (since the first subject randomized) is determined by reaching the specified sample size or event count
- Multiple analyses can be conducted for the same endpoint at different information levels

**Column Definitions:**

- **`endpoint`**: The endpoint that triggers the analysis (e.g., "CR", "OS", "EFS")
- **`strata`**: List of patient strata (or overall population) that constitute the sample size (or number of events) that triggers the analysis
- **`treatments`**: List of treatment arms whose patients contribute to the sample size (or number of events) that triggers the analysis. Note: In multi-armed studies, an analysis can be triggered by monitoring a subset of treatment arms
- **`sample_size`**: Target sample size for binary endpoints (NA for time-to-event)
- **`events`**: Target number of events for time-to-event endpoints (NA for binary)
- **`power_subsets_any`**: Named list of hypothesis subsets for "at least one rejection" power calculations. Each element is a vector of hypothesis indices (e.g., `list("H1, H2" = c(1, 2))` means power to reject at least one of H1 or H2)
- **`power_subsets_all`**: Named list of hypothesis subsets for "all rejections" power calculations. Each element is a vector of hypothesis indices (e.g., `list("H1, H2, H3" = c(1, 2, 3))` means power to reject all of H1, H2, and H3)



```{r analyses-detail}
# Show the structure of list columns
cat("Strata for Analysis 1:", paste(config$analyses$strata[[1]], collapse = ", "), "\n")
cat("Treatments for Analysis 1:", paste(config$analyses$treatments[[1]], collapse = ", "), "\n")
cat("\nPower subsets (any) for Analysis 1:\n")
print(config$analyses$power_subsets_any[[1]])
cat("\nPower subsets (all) for Analysis 1:\n")
print(config$analyses$power_subsets_all[[1]])
```

## 3. Hypotheses Definition

The `hypotheses` data frame specifies the statistical hypotheses to be tested. We display all columns in parts:

```{r hypotheses-part1}
# Part 1: Basic hypothesis definition
kable(config$hypotheses[,c("type", "endpoint", "strata", "control", "test")],
      caption = "Hypotheses - Part 1: Basic Definition")
```

```{r hypotheses-part2}
# Part 2: Analysis assignments
analyses_info <- data.frame(
  endpoint = config$hypotheses$endpoint,
  analyses_analysed = sapply(config$hypotheses$analyses_analysed,
                             function(x) paste(x, collapse = ", "))
)
kable(analyses_info, caption = "Hypotheses - Part 2: Analysis Assignments")
```

```{r hypotheses-part3}
# Part 3: Spending function details
spending_info <- config$hypotheses %>%
  select(endpoint, sf, sfpar, nominal) %>%
  mutate(
    sfpar = sapply(sfpar, function(x) if(is.null(x)) "NULL" else as.character(x)),
    nominal = sapply(nominal, function(x) if(is.null(x)) "NULL" else paste(x, collapse=", "))
  )
kable(spending_info, caption = "Hypotheses - Part 3: Spending Functions")
```

```{r hypotheses-part4}
# Part 4: Test method
test_info <- config$hypotheses %>%
  select(endpoint, test_method)
kable(test_info, caption = "Hypotheses - Part 4: Test Methods")
```

**Column Definitions:**

- **`type`**: Hypothesis type ("Primary" or "Secondary")
- **`endpoint`**: Endpoint being tested
- **`strata`**: List of strata for this hypothesis
- **`control`**: Control treatment arm name
- **`test`**: Test treatment arm name
- **`analyses_analysed`**: List or single value specifying which analyses (by row index in the `analyses` data frame) test this hypothesis
- **`sf`**: Spending function type ("none", "asHSD", "asOF", "asP", "asKD", "asUser") for group sequential design
- **`sfpar`**: Spending function parameter (e.g., gamma for HSD; NULL if not applicable)
- **`nominal`**: Nominal alpha spending at interim analyses (optional; NULL if not specified)
- **`test_method`**: Statistical test method ("logrank", "stratified_logrank", "unpooled_proportions", "pooled_proportions", "cmh")

**Spending Function Types:**
- `"none"`: No group sequential testing (single analysis)
- `"asHSD"`: Hwang-Shih-DeCani spending function
- `"asOF"`: O'Brien-Fleming spending function
- `"asP"`: Pocock spending function
- `"asKD"`: Kim-DeMets spending function
- `"asUser"`: User-defined spending

**Test Method Types:**
- `"logrank"`: Log-rank test for time-to-event endpoints (unstratified)
- `"stratified_logrank"`: Stratified log-rank test for time-to-event endpoints
- `"unpooled_proportions"`: Two-sample test for proportions (unpooled variance)
- `"pooled_proportions"`: Two-sample test for proportions (pooled variance)
- `"cmh"`: Cochran-Mantel-Haenszel test for stratified binary data

```{r hypotheses-detail}
# Show which analyses test each hypothesis
for(i in seq_len(nrow(config$hypotheses))) {
  cat("Hypothesis", i, "(", config$hypotheses$endpoint[i], "):")
  cat(" Analyses", paste(config$hypotheses$analyses_analysed[[i]], collapse = ", "), "\n")
}
```

## 4. Enrollment Rates

The `enroll_rate` data frame specifies patient enrollment assumptions:

```{r enrollment}
kable(config$enroll_rate, caption = "Enrollment Rate Specification")
```

**Column Definitions:**

- **`stratum`**: Patient stratum identifier
- **`treatments`**: List of treatment arms for this stratum
- **`rate`**: Enrollment rate (patients per month) for this stratum
- **`duration`**: Enrollment duration (months) for this stratum
- **`ratio`**: Randomization ratio vector for treatment arms (e.g., c(1, 1) for 1:1 randomization)

**Row Interpretation:**
- Each row represents enrollment for one stratum
- Total enrollment = rate × duration for each stratum
- Treatments list shows which arms patients in this stratum can be randomized to

```{r enrollment-detail}
# Calculate total enrollment
total_enrollment <- sum(config$enroll_rate$rate * config$enroll_rate$duration)
cat("Total Planned Enrollment:", total_enrollment, "patients\n")

# Show treatment allocation
for(i in seq_len(nrow(config$enroll_rate))) {
  cat("Stratum", config$enroll_rate$stratum[i], "treatments:")
  cat(paste(config$enroll_rate$treatments[[i]], collapse = ", "), "\n")
}
```

## 5. Time-to-Event Distribution Parameters

The `distribution_tte` data frame specifies parameters for time-to-event endpoints:

```{r tte-dist}
kable(config$distribution_tte, caption = "Time-to-Event Distribution Parameters", digits = 4)
```

**Column Definitions:**

- **`endpoint`**: Time-to-event endpoint name
- **`stratum`**: Patient stratum
- **`treatment`**: Treatment arm
- **`duration`**: Duration of constant hazard period (Inf = constant throughout)
- **`fail_rate`**: Hazard rate (events per month) for this period
- **`dropout_rate`**: Dropout hazard rate (per month)

**Row Interpretation:**
- Each row defines hazard rates for one stratum-treatment-endpoint combination
- Multiple rows per combination allow for piecewise constant hazards
- Dropout is assumed constant across all periods

```{r tte-detail}
# Calculate hazard ratios
tte_summary <- config$distribution_tte %>%
  filter(endpoint == "OS") %>%
  select(stratum, treatment, fail_rate) %>%
  tidyr::pivot_wider(names_from = treatment, values_from = fail_rate) %>%
  mutate(HR = TRT / PBO)

kable(tte_summary, caption = "Hazard Ratios for OS Endpoint", digits = 3)
```

## 6. Binary Endpoint Parameters

The `distribution_bin` data frame specifies binary endpoint parameters:

```{r bin-dist}
kable(config$distribution_bin, caption = "Binary Endpoint Parameters", digits = 3)
```

**Column Definitions:**

- **`endpoint`**: Binary endpoint name
- **`stratum`**: Patient stratum
- **`treatment`**: Treatment arm
- **`rate`**: Response rate (probability of success)
- **`maturity_time`**: Time (months) when endpoint can be evaluated

**Row Interpretation:**
- Each row defines response rate for one stratum-treatment-endpoint combination
- Maturity time determines when patients contribute to the analysis
- All patients must be followed for at least maturity_time months

```{r bin-detail}
# Calculate odds ratios
bin_summary <- config$distribution_bin %>%
  select(stratum, treatment, rate) %>%
  tidyr::pivot_wider(names_from = treatment, values_from = rate) %>%
  mutate(
    OR = (TRT / (1 - TRT)) / (PBO / (1 - PBO)),
    Risk_Diff = TRT - PBO
  )

kable(bin_summary, caption = "Treatment Effects for Binary Endpoints", digits = 3)
```

## 7. Graphical Testing Procedure

The `graph` component defines the multiple testing strategy:

```{r graph}
cat("Transition Matrix:\n")
print(config$graph$g)
cat("\nInitial Weights:\n")
print(config$graph$w)
```

**Components:**

- **`g`**: Transition matrix (3×3 for 3 hypotheses)
  - Element `g[i,j]` = fraction of alpha from hypothesis i transferred to hypothesis j when i is rejected
  - Diagonal elements should be 0
  - Row sums should be ≤ 1

- **`w`**: Initial weight vector
  - Element `w[i]` = initial alpha allocation to hypothesis i
  - Sum should equal 1.0
  - Determines initial local significance levels

```{r graph-detail}
# Verify graph properties
cat("Row sums of transition matrix:", rowSums(config$graph$g), "\n")
cat("Sum of initial weights:", sum(config$graph$w), "\n")
cat("Initial local alpha levels:", config$alpha * config$graph$w, "\n")
```

## Configuration Validation

The package includes validation to ensure configurations are properly specified:

```{r validation}
# Validate the configuration
is_valid <- validate_config(config)
cat("Configuration is valid:", is_valid, "\n")
```

## Processing the Configuration

Once configured, the study can be processed to generate all analysis components:

```{r processing}
# Process the configuration
result <- process_config(config)

# Show what gets generated
cat("Generated components:\n")
cat(paste("-", names(result), collapse = "\n"), "\n")
```

## Summary

A complete appendMCP configuration requires:

1. **Study metadata**: Name, description, alpha level
2. **Analyses schedule**: When analyses occur (by sample size or events), with power subset specifications for simulation-based operating characteristics
3. **Hypotheses**: What gets tested, how (spending functions), and which test method to use
4. **Enrollment**: Patient accrual rates by stratum with randomization ratios
5. **TTE distributions**: Survival parameters by stratum/treatment
6. **Binary distributions**: Response rates by stratum/treatment
7. **Graph structure**: Multiple testing procedure definition

Each component must be internally consistent and align with the others. The `validate_config()` function helps ensure proper specification before analysis.

## Best Practices

- **Start simple**: Begin with basic designs and add complexity gradually
- **Validate early**: Use `validate_config()` frequently during development
- **Document assumptions**: Clearly specify all distributional assumptions
- **Check consistency**: Ensure enrollment, analyses, and hypotheses align
- **Test scenarios**: Try different parameter values to understand sensitivity

For more examples and advanced configurations, see the other package vignettes and documentation.
