---
title: "Example Study Report - Group Sequential Design with Multiple Testing"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Example Study Report - Group Sequential Design with Multiple Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
# Time limit in plot
Tmax_atPlot = 60
# Number of rounding digits for sample sizes and events
n_digits = 0
# Number of digits to report for p-value boundaries
pval_digits = 5
# Number of rounding digits for information fractions
info_fraction_digits = 2
# Number of rounding digits for time
time_digits = 1
# Number of rounding digits for hurdle
hurdle_digits = 3
# Number of rounding digits for power
power_digits = 3
```

```{r setup, include=FALSE}
library(appendMCP)
library(knitr)

knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 6.5,
  fig.height = 4.5,
  message    = FALSE,
  warning    = FALSE,
  echo       = FALSE
)
options(width = 58)

config  <- load_config_from_repository("example_study")
results <- process_config(config)
```

## Introduction

With a `r if(!is.null(config$enroll_rate)) max(tapply(config$enroll_rate$duration, config$enroll_rate$stratum, sum)) else "XX"`-month accrual period, the total sample size planned for the study is `r if(!is.null(config$enroll_rate)) sum(config$enroll_rate$duration * config$enroll_rate$rate) else "XXX"`.

## Enrollment {.tabset}

### Piece-wise Constant Enrollment Rates

```{r enrollment_er_figure}
plot(results$er_figure)
```

### Cumulative Enrollment

```{r enrollment_er_cum_figure}
plot(results$er_cum_figure)
```

## Multiplicity Adjustment

The multiplicity strategy follows the graphical approach for group sequential designs of Maurer and Bretz (2013) which provides strong control of type 1 error. The procedure takes into account both sources of multiplicity: multiple hypothesis tests (e.g., across primary and secondary endpoints) and multiple analyses planned for the study (i.e., interim and final analyses).

There are two key components that define this approach:

- Testing algorithm for multiple hypotheses specified by the graphical representation
- Repeated testing of some hypotheses using the alpha-spending function methodology

The multiplicity strategy will be applied to the `r length(unique(results$hypotheses$index))` hypotheses across `r paste(unique(results$hypotheses$endpoint), collapse = ", ")` endpoints. The following table summarizes the hypotheses specifying alpha-spending functions (for hypotheses to be tested group sequentially) together with the effect sizes and planned maximum statistical information (sample size or number of events).

**Table 1. Summary of Primary and Key Secondary Hypotheses**

```{r table1, include=TRUE, echo=FALSE, results='markup'}
table1_hux(results$tables$table1, n_digits = n_digits)
```

The overall type I family-wise error rate for `r if(!is.null(config$graph$w)) length(config$graph$w) else "X"` hypotheses, over all (interim and final) analyses, is controlled to `r if(!is.null(config$alpha)) config$alpha*100 else 2.5`% (one-sided).

## Distribution Assumptions {.tabset}

### Binary Endpoints {.tabset}

#### Rates

```{r bin_figure}
if (!is.null(config$distribution_bin) && nrow(config$distribution_bin) > 0) {
  plot(results$bin_figure)
} else {
  cat("No binary endpoints specified.")
}
```

#### Absolute Risk Differences

```{r bin_rd_figure}
if (!is.null(config$distribution_bin) && nrow(config$distribution_bin) > 0) {
  plot(results$bin_rd_figure)
} else {
  cat("No binary endpoints specified.")
}
```

### Time-to-Event Endpoints {.tabset}

#### Drop-out Rates

```{r tte_dropout_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_dropout_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Cumulative Drop-out Probabilities

```{r tte_dropout_probability_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_dropout_probability_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Hazard Rate

```{r tte_hazard_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_hazard_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Cumulative Hazard

```{r tte_cumhaz_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_cumhaz_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Survival Curves

```{r tte_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Median Survival

```{r tte_median_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_median_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Survival Quantiles

```{r tte_quantiles_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_quantiles_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Enrollment Weighted Survival Curves

```{r tte_weighted_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_weighted_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Piece-wise Hazard Ratios

```{r tte_hr_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_hr_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

#### Average Hazard Ratios

```{r tte_ahr_figure}
if (nrow(config$distribution_tte) > 0) {
  plot(results$tte_ahr_figure)
} else {
  cat("No time-to-event endpoints specified.")
}
```

Figure 1 shows the graph where the hypotheses of interest are represented by the elliptical nodes. Each node has the hypothesis weight assigned to it (denoted by $w$). A particular value of $w$ sets the local significance level associated with that hypothesis (which is equal to `r if(!is.null(config$alpha)) config$alpha else 0.025` × $w$). The graphical approach allows local significance levels to be recycled (along arrows on the graph) when a given hypothesis is successful (i.e., the corresponding null hypothesis is rejected) at interim or final analyses.

**Figure 1. Graph Depicting Multiple Hypothesis Testing Strategy**

```{r MTgraph, echo=FALSE, fig.align="center", fig.height=4, fig.width=6}
plot(results$graph_figure)
```

## Interim Analyses {.tabset}

### IA by Hypothesis

**Table 2. Summary of Interim Analyses (by hypotheses)**

```{r iaDetailsTableA, include=TRUE, echo=FALSE, results='markup'}
table2_hux(results$tables$table2, n_digits = n_digits,
           time_digits = time_digits, info_fraction_digits = info_fraction_digits)
```

### IA by Calendar Time

**Table 3. Summary of Interim Analyses (by calendar analysis)**

```{r iaDetailsTableB, include=TRUE, echo=FALSE, results='markup'}
table3_hux(results$tables$table3, n_digits = n_digits,
           time_digits = time_digits, info_fraction_digits = info_fraction_digits)
```

### Information Flow

**Figure 2. Information Factor Over Time by Hypothesis**

```{r informationPlot, echo=FALSE, fig.align="center", fig.height=6, fig.width=9}
plot(results$information_figure)
```

### Timeline of Planned Analyses

**Figure 3. Anticipated study timeline**

```{r timelineType1Plot, echo=FALSE, fig.align="center", fig.height=3, fig.width=9}
plot(results$timeline_type1_figure)
```

**Figure 3b. Anticipated time of each analysis, with associated trigger**

```{r timelineType2Plot, echo=FALSE, fig.width=10, fig.height=4}
plot(results$timeline_type2_figure)
```

## Hypothesis Testing {.tabset}

### Scenarios

**Table 4. Weight Allocation Scenarios**

```{r graphTable, include=TRUE, echo=FALSE, results='markup'}
table4_hux(results$tables$table4, pval_digits = pval_digits)
```

### Boundary Specifications

**Table 5. Boundary Specifications**

Table 5 details the hypothesis testing at the interim and final analyses. For hypotheses tested group sequentially, the table provides the nominal p-value boundary derived from the alpha-spending function and the information fractions. The timing of analyses is expressed in terms of statistical information fractions. The table also reports local power at each analysis time.

```{r grSeqTable, include=TRUE, echo=FALSE, results='markup'}
table5_hux(results$tables$table5, info_fraction_digits = info_fraction_digits,
           pval_digits = pval_digits, hurdle_digits = hurdle_digits,
           prob_digits = power_digits)
```

### Operating Characteristics by Analysis

**Table 6a. Operating Characteristics at Each Analysis**

```{r operatingCharByAnalysis, include=TRUE, echo=FALSE, results='markup'}
if (!is.null(results$tables$table6a)) {
  table6a_hux(results$tables$table6a, prob_digits = power_digits)
} else {
  cat("Simulation results not available.")
}
```

### Operating Characteristics Overall

**Table 6b. Operating Characteristics Across Analyses**

```{r operatingCharAcrossAnalyses, include=TRUE, echo=FALSE, results='markup'}
if (!is.null(results$tables$table6b)) {
  table6b_hux(results$tables$table6b)
} else {
  cat("Simulation results not available.")
}
```

### Spending Functions

**Figure 4. Alpha-spending functions**

```{r alphaSpendingPlot, eval=TRUE, echo=FALSE, fig.height=8, fig.width=10}
plot(results$alpha_spend_figure)
```

## Configuration Details

### Graph Structure

**Table 7. Transition Matrix**

```{r graph-matrix}
kable(config$graph$g, digits = 3)
```

**Initial Weights:** `r paste(config$graph$w, collapse = ", ")`

### Enrollment Assumptions

**Table 8. Enrollment Rate Assumptions**

```{r enrollment_input}
kable(config$enroll_rate, digits = 2)
```

### Binary Endpoint Assumptions

**Table 9. Binary Endpoint Parameters**

```{r bin_input}
if (!is.null(config$distribution_bin) && nrow(config$distribution_bin) > 0) {
  kable(config$distribution_bin, digits = 3)
} else {
  cat("No binary endpoints specified.")
}
```

### Time-to-Event Endpoint Assumptions

**Table 10. Time-to-Event Distribution Parameters**

```{r tte_input}
if (!is.null(config$distribution_tte) && nrow(config$distribution_tte) > 0) {
  kable(config$distribution_tte, digits = 4)
} else {
  cat("No time-to-event endpoints specified.")
}
```

## References

Maurer W, Bretz F. Multiple testing in group sequential trials using graphical approaches. *Statistics in Biopharmaceutical Research.* 2013;5(4):311-320.

---

*Report generated on `r Sys.time()`*
