---
title: "Platform Trial with Arm Added Midway"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Platform Trial with Arm Added Midway}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE}
n_digits             <- 0
pval_digits          <- 5
info_fraction_digits <- 1
time_digits          <- 1
hurdle_digits        <- 3
prob_digits          <- 1
analysis_digits      <- 2
weight_digits        <- 2
```

```{r setup, include=FALSE}
library(appendMCP)
library(knitr)

knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 6.5,
  fig.height = 4.5,
  message    = FALSE,
  warning    = FALSE,
  echo       = FALSE
)
options(width = 58)

config  <- load_config_from_repository("platform_study")
results <- process_config(config)
```

## Introduction

This vignette demonstrates how `appendMCP` handles a **platform-type group sequential design** in which a new treatment arm is added midway through an ongoing trial.
The design pattern illustrated here extends the standard two-arm group sequential framework described in the `example_study_report` vignette to accommodate:

- A **3-arm structure** (TrtA monotherapy, TrtAB combination, Control) where TrtAB is added after the trial has already started enrolling TrtA vs Control;
- A **biomarker stratification** (B+ / B−) that is a genuine patient subgroup of clinical interest;
- **Two control-arm pooling strategies** — pooled and concurrent — for the TrtAB vs Control comparison;
- A **large graphical multiple testing procedure** spanning 12 hypotheses across 5 endpoints.

The study has a total one-sided family-wise type I error rate of `r config$alpha * 100`% distributed across `r sum(config$graph$w > 0)` primary hypotheses, each receiving an initial weight of `r unique(config$graph$w[config$graph$w > 0])`.
The inflated `alpha = 0.025 × 5` is an accounting device: it allows `process_config()` to compute power for all 12 hypotheses (including secondary endpoints with no formal alpha allocation), while each primary hypothesis still operates at a local significance level of 0.025.

## Study Design

### Two-Part Enrollment

The trial enrolls in two parts:

- **Part 1** — TrtA vs Control (2:1 ratio), B+/B− stratified, over approximately 14 months.
- **Part 2** — TrtAB arm opens. All three arms enroll (TrtA : Control : TrtAB = 1:1:4 ratio). B+/B− stratification continues.

The total accrual period is `r max(tapply(config$enroll_rate$duration, config$enroll_rate$stratum, sum))` months.

### The Four-Stratum Structure

`appendMCP` uses a **stratum** field to define subpopulations for enrollment and event modeling.
This study uses 4 strata that encode two independent dimensions simultaneously:

| Stratum | Enrollment stage | Biomarker |
|---|---|---|
| `S1_Bpos` | Part 1 (before TrtAB opens) | B+ |
| `S1_Bneg` | Part 1 (before TrtAB opens) | B− |
| `S2_Bpos` | Part 2 (after TrtAB opens)  | B+ |
| `S2_Bneg` | Part 2 (after TrtAB opens)  | B− |

The S1/S2 dimension is **artificial** — it is introduced purely to track which patients were enrolled before vs after the TrtAB arm opened.
This is necessary because the two groups of control arm patients may not be exchangeable (a perceived stage effect), which motivates the concurrent control analysis.
The B+/B− dimension is a **real** stratification factor that defines a subgroup of primary clinical interest (H12).

### Encoding "Arm Added Midway" via Zero Enrollment Rates

The platform design is encoded in the `enroll_rate` table using **zero-rate placeholder rows**.
S2 strata have rate = 0 during Part 1 (months 0–14), signaling that no S2 patients exist yet.
S1 strata have no rows after Part 1, so their enrollment stops when Part 1 ends.

```{r enroll-table, echo=FALSE}
kable(config$enroll_rate[, c("stratum", "treatments", "rate", "duration", "ratio")],
      caption = "Enrollment rate specification. S2 rows with rate = 0 are required placeholders during Part 1.",
      digits = 1)
```

The enrollment curves below show the resulting piecewise accrual by arm and stage.

## Enrollment {.tabset}

### Piece-wise Constant Enrollment Rates

```{r er_figure}
plot(results$er_figure)
```

### Cumulative Enrollment

```{r er_cum_figure}
plot(results$er_cum_figure)
```

## Multiplicity Adjustment

The multiplicity strategy follows the graphical approach of Maurer and Bretz (2013), controlling the family-wise type I error rate across both multiple hypotheses and multiple analyses.

The `r length(unique(results$hypotheses$index))` hypotheses span `r paste(unique(results$hypotheses$endpoint), collapse = ", ")` endpoints.
The following table summarizes the hypotheses with their alpha-spending functions and maximum statistical information.

**Table 1. Summary of Hypotheses**

```{r table1, results='markup'}
table1_hux(results$tables$table1, n_digits = n_digits)
```

### Pooled vs Concurrent Control

For the TrtAB vs Control comparison, two separate sets of hypotheses are pre-specified:

- **Pooled** (H3–H7): Control arm patients from both Part 1 and Part 2 are combined. This maximizes power but assumes no stage effect on the control arm.
- **Concurrent** (H8–H9): Only control arm patients enrolled concurrently with TrtAB (Part 2, S2 strata) are used. This is conservative but robust to a potential stage effect.

Both strategies are included in the graphical testing procedure, with alpha propagating between them upon rejection.

### Alpha Distribution

The overall family-wise type I error rate is controlled at 2.5% (one-sided) per primary hypothesis.
The `alpha = 0.025 × 5` setting in the config is an accounting device: with 5 primary hypotheses each receiving initial weight `r unique(config$graph$w[config$graph$w > 0])`, each operates at a local significance level of `r unique(config$graph$w[config$graph$w > 0])` × `r config$alpha` = 0.025.
Secondary hypotheses (H2, H4, H5) have initial weight 0 and receive alpha only through propagation upon rejection of primary hypotheses.

**Figure 1. Graphical Multiple Testing Procedure**

```{r MTgraph, fig.align="center", fig.height=5, fig.width=6}
plot(results$graph_figure)
```

## Interim Analyses {.tabset}

Three event-driven analyses are planned:

1. **IA1** — `r config$analyses$events[1]` OS events in S1 strata (TrtA vs Control)
2. **IA2** — `r config$analyses$events[2]` OS events in S1 strata (TrtA vs Control, final for Part 1) and first look at TrtAB vs Control
3. **FA** — `r config$analyses$events[3]` OS events in S1+S2 strata (TrtAB vs Control, final)

### IA by Hypothesis

**Table 2. Summary of Interim Analyses (by hypothesis)**

```{r iaDetailsTableA, results='markup'}
table2_hux(results$tables$table2,
           n_digits             = n_digits,
           time_digits          = time_digits,
           info_fraction_digits = info_fraction_digits)
```

### IA by Calendar Time

**Table 3. Summary of Interim Analyses (by calendar analysis)**

```{r iaDetailsTableB, results='markup'}
table3_hux(results$tables$table3,
           n_digits             = n_digits,
           time_digits          = time_digits,
           info_fraction_digits = info_fraction_digits)
```

### Information Timeline

```{r informationPlot, fig.align="center", fig.height=6, fig.width=9}
plot(results$information_figure)
```

### Timeline: Overall

```{r timelineType1Plot, fig.align="center", fig.height=5, fig.width=9}
plot(results$timeline_type1_figure)
```

### Timeline: Triggers

```{r timelineType2Plot, fig.width=10, fig.height=4}
plot(results$timeline_type2_figure)
```

### Spending Functions

```{r alphaSpendingPlot, fig.height=8, fig.width=10}
plot(results$alpha_spend_figure)
```

## Hypothesis Testing {.tabset}

### Weight Scenarios

**Table 4. Weight Allocation Scenarios**

```{r graphTable, results='markup'}
table4_hux(results$tables$table4, pval_digits = pval_digits)
```

### Boundary Specifications

**Table 5. Boundary Specifications**

```{r grSeqTable, results='markup'}
table5_hux(results$tables$table5,
           info_fraction_digits = info_fraction_digits,
           pval_digits          = pval_digits,
           hurdle_digits        = hurdle_digits,
           prob_digits          = prob_digits)
```

## Operating Characteristics {.tabset}

### By Analysis

**Table 6a. Operating Characteristics at Each Analysis**

```{r operatingCharByAnalysis, results='markup'}
if (!is.null(results$tables$table6a)) {
  table6a_hux(results$tables$table6a, prob_digits = prob_digits)
} else {
  cat("Simulation results not available.")
}
```

### Across Analyses

**Table 6b. Operating Characteristics Across Analyses**

```{r operatingCharAcrossAnalyses, results='markup'}
if (!is.null(results$tables$table6b)) {
  table6b_hux(results$tables$table6b,
              analysis_digits = analysis_digits,
              time_digits     = time_digits)
} else {
  cat("Simulation results not available.")
}
```

## Distribution Assumptions {.tabset}

### Time-to-Event Endpoints {.tabset}

#### Survival Curves

```{r tte_figure}
plot(results$tte_figure)
```

#### Hazard Rates

```{r tte_hazard_figure}
plot(results$tte_hazard_figure)
```

#### Hazard Ratios

```{r tte_hr_figure}
plot(results$tte_hr_figure)
```

#### Average Hazard Ratios

```{r tte_ahr_figure}
plot(results$tte_ahr_figure)
```

#### Median Survival

```{r tte_median_figure}
plot(results$tte_median_figure)
```

#### Drop-out Rates

```{r tte_dropout_figure}
plot(results$tte_dropout_figure)
```

## Input Data {.tabset}

### Enrollment

```{r enrollment_input}
kable(config$enroll_rate, caption = "Enrollment Rate Assumptions", digits = 2)
```

### Time-to-Event Distributions

```{r tte_input}
kable(config$distribution_tte, caption = "Time-to-Event Distribution Parameters", digits = 4)
```

### Graph Structure

**Initial weights:** `r paste(round(config$graph$w, weight_digits), collapse = ", ")`

```{r graph-matrix}
kable(config$graph$g, caption = "Transition Matrix", digits = 3)
```

## References

Maurer W, Bretz F. Multiple testing in group sequential trials using graphical approaches. *Statistics in Biopharmaceutical Research.* 2013;5(4):311-320.
